Testing the social, cognitive, and motor foundations of Paleolithic skill reproduction

### 2022-05-10 manuscript/manuscript.Rmd
made a few minor changes to figures.

**Table** \@ref(tab:tab1) lists details for the study’s various lithic attributes. We weighed and measured all the finished cores (L, W, T). Delta weight was calculated as (nodule start weight-core end weight)/nodule start weight. All detached pieces (DPs) were collected and weighed. We did not sort DPs into types (e.g., whole flakes, fragments), which would have greatly increased processing time. It is not clear that such distinctions add relevant information regarding utility/desirability beyond that supplied by metrics [@stout2019]. All DPs larger than 40mm in maximum dimension were photographed and measured. It is conventional in Early Stone Age lithic analysis to employ a 20 mm cut-off. We selected a higher threshold for both pragmatic (analysis time) and theoretical reasons. Flake use experiments have shown that flakes weighing less than  5–10 g or with a surface area below 7–10 $cm^2$ [@prasciunas2007] or with a maximum dimension \<50-60 mm [@key2014] become markedly inefficient for basic cutting tasks. Similarly, data from Oldowan replication experiments [@stout2019] show that the utility index (flake cutting edge/flake mass$^{1/3}$) \* (1 - exp[-0.31 \* (flake maximum dimension - 1.81)]) developed by Morgan et al. (2015) falls off rapidly below 40mm maximum dimension (Mean Utility \< 40mm = 0.508; \>=40mm = 0.946; t= 11.99, df = 707, p \< 0.000). By including weight in our cut-off criteria, we also avoid skewing the flake shape distribution by selectively retaining long, thin pieces (i.e., MD \> 40, weight \< 5g) while discarding rounder pieces of similar (or greater) weight and area. We recorded the absolute number of above-threshold DPs produced per core and calculated a mass-based proportion of larger DPs by dividing the combined weight of DPs \>40mm in maximum dimension and 5g in weight by the weight of all DPs. Higher values show cores from which proportionally more of the detached mass comprises large pieces.
```{r tab1, echo=FALSE, message=FALSE}
library(tidyverse)
df1 <- data.frame("Variable" = c("Delta mass", "Mass of large detached pieces / total debitage mass", "Length", "Width", "Thickness"), 
                 "Variable class" = c("Ratio", "Ratio", "Linear measurement", "Linear measurement", "Linear measurement"), 
                 "Definition" = c("Final core mass as a percentage of original nodule mass. Higher values show more completely flaked nodules.", "Combined mass of all detached pieces >40mm in maximum dimension and 5g in mass divided by the mass of all material detached from a specific core. Higher values show cores with more detached pieces per knapped material.", "Measurement along each core and flake's longest axis in mm.", "Measurement taken at 10% increments along a detached piece's longest axis starting at the detached piece's tip (in mm). Same measurement taken orthogonal to a core's length.", "Measurement taken at the maximum point in a detached piece's profile and orthogonal to width on the cores."),
                "Recording method" = c("Digital scale", "Calipers, counts, digital scale", "2D photogrammetry", "2D photogrammetry", "Calipers") )

>>>>

**Table** \@ref(tab:tab1) lists details for the study’s various lithic attributes. We weighed and measured all the finished cores (L, W, T). Delta weight was calculated as (nodule start weight-core end weight)/nodule start weight. All detached pieces (DPs) were collected and weighed. We did not sort DPs into types (e.g., whole flakes, fragments), which would have greatly increased processing time. It is not clear that such distinctions add relevant information regarding utility/desirability beyond that supplied by metrics [@stout2019]. All DPs larger than 40mm in maximum dimension were photographed and measured. It is conventional in Early Stone Age lithic analysis to employ a 20 mm cut-off. We selected a higher threshold for both pragmatic (analysis time) and theoretical reasons. Flake use experiments have shown that flakes weighing less than  5–10 g or with a surface area below 7–10 $cm^2$ [@prasciunas2007] or with a maximum dimension \<50-60 mm [@key2014] become markedly inefficient for basic cutting tasks. Similarly, data from Oldowan replication experiments [@stout2019] show that the utility index (flake cutting edge/flake mass$^{1/3}$) \* (1 - exp[-0.31 \* (flake maximum dimension - 1.81)]) developed by Morgan et al. (2015) falls off rapidly below 40mm maximum dimension (Mean Utility \< 40mm = 0.508; \>=40mm = 0.946; t= 11.99, df = 707, p \< 0.000). By including weight in our cut-off criteria, we also avoid skewing the flake shape distribution by selectively retaining long, thin pieces (i.e., MD \> 40mm, weight \< 5g) while discarding rounder pieces of similar (or greater) weight and area. We recorded the absolute number of above-threshold DPs produced per core and calculated a mass-based proportion of larger DPs by dividing the combined weight of DPs \>40mm in maximum dimension and 5g in weight by the weight of all DPs. Higher values show cores from which proportionally more of the detached mass comprises large pieces.
```{r tab1, echo=FALSE, message=FALSE}
library(tidyverse)
df1 <- data.frame("Variable" = c("Delta mass", "Mass of large detached pieces / total debitage mass", "Length", "Width", "Thickness"), 
                 "Variable class" = c("Ratio", "Ratio", "Linear measurement", "Linear measurement", "Linear measurement"), 
                 "Definition" = c("Final core mass as a percentage of original nodule mass. Higher values show more completely flaked nodules.", "Combined mass of all detached pieces >40mm in maximum dimension and 5g in mass divided by the mass of all material detached from a specific core. Higher values show cores with more detached pieces per knapped material.", "Measurement along each core and flake's longest axis in mm.", "Measurement taken at 10% increments along a detached piece's longest axis starting at the detached piece's tip (in mm). Same measurement taken orthogonal to a core's length.", "Measurement taken at the maximum point in a detached piece's profile and orthogonal to width on the cores."),
                "Recording method" = c("Digital scale", "Calipers, counts, digital scale", "2D photogrammetry", "2D photogrammetry", "Calipers") )
